Individual and Differential Harm in Redistricting

### 2022-05-06 paper/paper.Rmd
move proof

This definition of harm is limited in two key regards.
First, it considers only individuals who vote; if a person abstains from voting then their calculated harm is zero.
A natural extension of this definition would be to separate an individual's candidate preference from their vote choice; the latter would be used in calculating election winners, but the former could provide a more broadly-based harm measure.
Naturally, this approach requires auxiliary data to implement (such as a large-scale survey), since candidate preferences for nonvoters cannot be inferred from election returns.

@@

Estimating these quantities from available data is another matter, of course, and requires additional assumptions, as we describe next.

# Estimating the Harm Measures {#sec:estimating}

@@

<!-- ## Simulation Methods -->

<!-- Despite being derived from individual harm, differential partisan harm is in fact closely related to the overall number of seats won by each party, as Proposition \ref{prop:fairness} describes below. -->

<!-- Yet unlike a simple measure of seats, differential harm is shifted to an interpretable scale, where a value of zero corresponds to fairness. -->

<!-- In other words, differential harm provides a connection between fairness and seats, establishing a normative guide to overall election outcomes. -->

<!-- This connection further supports the use of total seats won by each party as a summary statistic in partisan redistricting simulation methods, as advocated by @deford2020symmetry. -->

```{=tex}
\begin{proprep} \label{prop:fairness}
Suppose $\Groups=\{\dem,\rep\}$. If turnout is equal across districts, then for a specific districting plan $q\in\Plans$,
\begin{align*}
    \Delta_{\rep-\dem} H(q) = \E[\seats(\dem, V, \tilde Q)] - \E[\seats(\dem, V, q)].
\end{align*}
\textbf{CM note: define $\seats$ or replace with sum.}
\end{proprep}
\begin{proof}
\end{proof}
```
# Application to Partisan Gerrymandering in New Jersey {#sec:nj}

>>>>

This definition of harm is limited in two key regards.
First, it considers only individuals who vote; if a person abstains from voting then their calculated harm is zero.
A natural extension of this definition would be to separate an individual's candidate preference from their vote choice; the latter would be used in calculating election winners, but the former could provide a more broadly-based harm measure.
Naturally, this approach requires auxiliary data to implement (such as a large-scale survey), since candidate preferences for nonvoters cannot be directly inferred from election returns.

@@

Estimating these quantities from available data is another matter, of course, and requires additional assumptions, as we describe in the next section.
First, however, we note two interesting consequences of our definition of differential harm.

In the partisan case, with $\Groups=\{\dem,\rep\}$, differential harm is in fact closely related to the overall number of seats won by each party, as Proposition \ref{prop:fairness} describes below.
Yet unlike a simple measure of seats, differential harm is shifted to an interpretable scale where a value of zero corresponds to fairness.
In other words, differential harm provides a connection between fairness and seats, establishing a normative guide to overall election outcomes.
This connection further supports the use of total seats won by each party as a summary statistic in partisan redistricting simulation methods, as advocated by @deford2020symmetry.

```{=tex}
\begin{proprep} \label{prop:fairness}
Suppose $\Groups=\{\dem,\rep\}$. If turnout is equal across districts, then for a specific districting plan $q\in\Plans$,
\begin{align*}
    \Delta_{\rep-\dem} H(q) = \E[\seats(\dem, V, \tilde Q)] - \E[\seats(\dem, V, q)].
\end{align*}
\textbf{CM note: define $\seats$ or replace with sum.}
\end{proprep}
\begin{proof}
\end{proof}
```
# Estimating the Harm Measures

@@

# Application to Partisan Gerrymandering in New Jersey
